Word Learning as Network Growth: A Cross-linguistic Analysis

### 2018-01-30 cogsci/cogsci.Rmd
modif to cogsci draft

---
title: "Word Learning as Network Growth: a Cross-linguistic Analysis"

@@

@@

    "Children tend to produce first the words that are connected to a variety of other words along both the phonological and the semantic dimensions. Though this connectivity effect has been extensively documented, little is known about the underlying developmental mechanism.  One view suggests that learning is primarily driven by a network growth model where highly connected words in the child's ealy lexicon attract similar words. Another view suggests that learning is driven, not by highly connected words in the early internal lexicon, but by highly connected word in the external learning environment. The present study tests both scenarios systematically in both the phonological and the semantic domains, and across 8 languages. We show that external connectivity in the learning environment drives growth in both the semantic and the phonological networks, and that this pattern is consistent cross-linguistically. The findings suggest a word learning mechanism where children harness their statistical learning abilities to (indirectly) detect and learn highly connected words in the learning environment."

@@

<!-- Children tend to learn some words typically at an earlier age than others. Researchers have documented many word properties that influence the order of acquisition of words.  This paper investigates strcutural properties that characterize how words are related to each other in the lexicon.  -->
```{r global_options, include=FALSE}
rm(list=ls())

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
options(xtable.comment = FALSE)
library(purrr)
library(readr)
library(ggplot2)
library(langcog)
library(boot)
#library(lazyeval)
library(dplyr)
library(tidyr)
library(wordbankr)
library(directlabels)
#library(scales)
library(stringr)
library(lmtest)
library(rwebppl)
library(jsonlite)
library(nlme)
library(feather)
library(broom)
library(HDInterval)
library(BBmisc)
```

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from controlled laboratory experiments to corpus analysis. One strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn. For example, within a lexical category, words that are more frequent in child-directed speech are acquired earlier [@goodman2008]. Other factors include word length, the mean length of utterances in which the word occurs, and how concrete the word is [see @braginsky2016].

Besides these word-level properties, researchers found that the lexical structure (that is, how words relate to each other) also influences the age of acquisition of words. The lexical structure is best characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words.  Previous studies investigated early vocabulary networks based on different word relations such as shared semantic features, target-cue relationship in free association norms, co-occurrence in child directed speech, and phonological similarity. These studies found that children tend to produce earlier the words that have higher neighborhood density (i.g., high connectivity in the network) both at the phonological and the semantic level [@stokes2010; @storkel2009; @hills2009; @hills2010; @engelthaler2017; @stella2017; @beckage2011].

While most studies focued on the static properties of the lexical network, a few have proposed to account for the underlying developmental process. @steyvers2005 suggested that the effect of connectivity is the consequence of how the lexical network gets constructed in the child's mind. According to this explanation, known as Preferential Attachment (PAT), highly connected words in the child's lexicon tend to "attract" more words over time, in a rich-get-richer scenario [@barabasi99]. In other words, what predicts word learning is *internal* connectivity in the child's early lexicon. In contrast, @hills2009 suggested that what biases the learning is not the connectivity in the child's internal lexicon but, rather, *external* connectivity in the learning environment. They called this alternative explanation Preferential Acquisition (PAC).  Figure \ref{fig:growth} shows an illustration of both growth scenarios with the same simplified network.

```{r growth, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=2, fig.cap = "\\label{fig:growth}Illustration of the two growth models with the same network. Filled circles (I1-I4) represent known words (internal lexicon), and empty circles (E1 and E2) represent words that have not been learned yet (external lexicon). Black lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant. For PAT, each candidate node is characterized with the average degree (i.e., number of links) of the existing nodes that it would attach to. Thus, according to PAT, the node E1 is more likely to enter the lexicon first. For PAC, each candidate node is characterized with its degree in the entire network. According th PAC, the node E2 is more likely to enter the lexicon first."}

@@

Studies that investigated lexical network growth focused on semantic networks using English data only [@steyvers2005; @hills2009; @hills2010].  The novelty of the current study is that it investigates whether phonological networks, like semantic networks, grow by PAC, or if they rather grow by PAT. It also provides a systematic comparison of both network growth scenarios in the phonological and the semantic domains and assesses their relative contribution to the learning process. Moreover, it tests the generality of the findings across eight languages.

The paper is organized as follows. First, we describe the dataset we used, the procedure we followed to construct both semantic and phonological networks. Next, we show the results of the various analyses we conducted comparing network growth scenarios across languages. Then we test the overall contribution of the networks in the learning process relative to other known predictors of age of acquisition (frequency and length). Finally we discuss the major results in relation to known experimental facts in language development, and we speculate about the potential learning mechanism that might drive early vocabulary learning.

@@

plot_data_all <- ggplot(data_long, aes(x=value, y=AoA))+
  facet_grid(predictor ~ language)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_bw()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=correlations, aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("AoA")

@@

plot_dist <- ggplot(data = degreeDist,  aes(x=x, y=y, col=dimension))+
  facet_grid(. ~ language)+
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme_bw()+
  theme(aspect.ratio = 0.7)+
  xlab("degree") +ylab("probability")

@@

# Networks
## Data
We used data from Wordbank [@frank2017], an open repository aggregating cross-linguistic language developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), a parent report vocabulary checklist. We used the *Words and Sentence* version of the CDI which contains the productive vocabulary of toddlers (age varied between 16 to 36 months). Following previous studies [e.g., @storkel2009; @hills2009], we restricted our analysis to the category of nouns. The age of acquisition was defined by the month at which a word was produced by at least 50% of children [@goodman2008].

We obtained these nouns in eight languages: Croatian, Danish, English, Italian, Norwegian, Russian, Spanish, and Turkish. We used the subset of nouns that had entries in the Florida Association Norms (see below). Since these norms are available only in English, we used the translation of non-English words provided by @braginsky2016. This allowed us to use the English association norms across languages. Table \ref{tab:stats} gives an overview of the data used. The translation of non-English words is still an ongoing process. Note, however, that all languages have at least 60% of nouns translated.

@@

tab <- xtable::xtable(statistics,
                      caption = "\\label{tab:stats}Total number of productive nouns in the CDI (left). We used a subset of these nouns that had available English translations (middle). The final set consisted of nouns that had both available translations as well entries in the Free Association Norms (right).")

@@

We used as an index of semantic relatedness the Florida Free Association Norms [@nelson1998]. This dataset was collected by giving adult participants a word (the cue), and asking them to write the first word that comes to mind (the target). For example, when given the word "ball", they might answer with the word "game". A pair of nodes were connected by a directed link from the cue to the target if there was a cue-target relationship between these nodes in the association norms. Each node was characterized by its "indegree", which represents the number of links for which the word was the target. To model growth from month to month, we constructed a different network at each month, made of words that have been learned at that month. Following @hills2009, we used as a proxy for the learning environment, the network constructed using the full set of nouns in the CDI data in each language (e.g., in English this corresponds to the network by 30 months).

@@

We used as a measure of phonological relatedness between two nodes the Levenshtein (edit) distance between their phonological forms. The measure counts the minimum number of operations (insertions, deletions, substitutions) required to change one string into another.  We generated approximate International Phonetic Alphabet (IPA) transcriptions from the orthographic transcription, across languages, using the open source text-to-speech software **[Espeak](http://http://espeak.sourceforge.net/).**

In previous studies, two nodes were linked if they had an edit distance of 1 [@stokes2010; @storkel2009]. However, in these previous studies the network was built using an adult vocabulary. Here, similar to the approximation we made for the semantic network, we used the full set of nouns in the CDI data as a proxy for phonological connectivity in the learning environment. However, since the children's vocabulary contains very few word pairs with an edit distance of 1, the resulting network was too sparse and uninformative. Thus, we increased the threshold from 1 to 2, that is, two node were related if their edit distance was equal to 1 or 2. Each node was characterized with its degree, i.e., the number of links it shares with other words. <!-- Moreover, and in order to model growth from month to month, we constructed a different network at each month, made of words that have been learned at that month. -->

@@

## Preliminary analysis
We start by examining how connectivity in the global network influences the age of acquisition. Regardless of the growth mechanism that might have given rise to this network (i.e., PAC or PAT), we expect nodes with the higher degrees in this network to be learned earlier than nodes with the lower degrees. The global network was constructed using nouns at the oldest age for which we have CDI data. Figure \ref{fig:data_all} shows how the age of acquisition for each word varies as a function of its degree (or indegree for the semantic network). For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled across languages. The plots show, overall, a negative correlation between the month of acquisition and the degree, indicating that nouns with higher degrees are generally learned earlier.

We also analysed the global network degree distribution. PAT is known to generate networks with a power-law degree distribution [@barabasi99]. If the networks display this property, this fact would strongly suggests a PAT-like generative process. <!--cannot be ruled out *a priori* as a potential growth mechanism. In contrast, if there is strong evidence against a power law distribution, then PAT (but not PAC) can be ruled out--> The results are shown in Figure \ref{fig:degree_distribution}. Overall, the extent to which the degree distribution approximates a power law varies across dimensions and languages. For a more direct test of the developmental process, we fit explicit growth models to the data (see next section).
<!--
```{r }

@@

To test the network growth scenarios, we fit different growth models to the data. We proceeded as follows. We calculated the probability that a word $w_i$, with a growth value $d_i$ would enter the lexicon at a given month, using a softmax function: 
\begin{equation}
 p(w_i)= \frac{e^{\beta d_i}}{\sum_j e^{\beta d_j} }
\end{equation}
Where $\beta$ is the fitting parameter. A positive value of $\beta$ means that words with higher growth values $d_i$ are acquired first, and a negative value means that words with lower growth values are acquired first. As explained in Figure \ref{fig:growth}, each candidate word $w_i$ is characterized with a growth value $d_i$ that depends on the growth scenario tested.  In PAT, the growth value of a given word is equal to the average degree of the existing nodes that it would attach to. In PAC, the growth value of a word is equal to its degree in the entire network.  The normalization includes all words that could be learned at that month.

We estimated the parameter $\beta$ using a Bayesian approach. The inference was performed using the probabilistic programming language  WebPPL [@dippl]. We started with a uniform distribution, and at each month, we computed the likelihood function over words that could possibly enter the lexicon at that month, fit to the words that have been actually learned at that month (using formula 1). We ended up with a posterior distribution of $\beta$, which we summarized in Figure \ref{fig:pred_ind}.

Besides fitting a growth model to the data, we conducted another evaluation. This second evaluation consists in determining, in each month, the model-dependent growth value distribution of all words that could possibly be learned at this month, and then computing the z-score of each learned word with respect to this distribution.  For each model, we tested if the distribution constituted by the z-scores of all learned words was different from zero, using a one-sample t-test.

The results from both evaluations were very similar and lead essentially to the same conclusions\footnote{we do not show here the results of the second evaluation because they were redundant with the results of the first evaluation}. For the semantic network, the results replicate Hills et al.'s finding in English, which was that the semantic network grows by PAC, not by PAT. Moreover, it generalizes this finding to all other languages (with the exception of Russian). For the phonological network, the results show that, generally speaking, PAC again fits the developmental trajectory better than PAT. We note however that PAT, though weaker, fares  better for the phonological networks (where it predicts part of the growth process in some languages such as Croatian, English, Norwegian and Russian) than it does for the semantic networks (where it is rather universally unpredictive).

@@

Above we evaluated the network growth scenarios individually. As a next step, we analysed their relative contribution to the learning process. This was done through adding more fitting parameters to the model, that is, by substituting $\beta d_i$ in formula (1) with: $$\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}$$ where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and PhonoPAC. Using the same fitting technique, we obtained the values shown in Figure \ref{fig:pred_ind}. In term of growth scenarios, we found that PAC dominates the learning. <!--In particular, though we found previously that phonological PAT predicted part of the learning in some languages when tested individually, here PAT appears to lose its predictive power when pitted against phonological PAC--> In terms of domain, both phonological and semantic networks appear to contribute to the lexical growth, but the phonological network appears to be stronger and more consistent across languages. In summary, the findings show that both semantic and phonological networks grow primarily by PAC, and that, generally speaking, semantic and phonological networks both contribute to the learning process.

@@

We saw that the way semantic and phonological information is structured in the learning environment (i.e., PAC) contribute to noun learning across languages. However, we know that other factors influence learning as well [e.g., @braginsky2016]. In what follows, we will investigate how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which correlates with phonological connectivity. <!-- [@pisoni1985].-->

Since PAT was uninformative, we dropped it from this analysis. Thus, we no longer needed to fit the growth model month-by-month as in the previous section. In fact, word utilities (i.e., growth values) in the case of PAC are fixed, they do not depend on previously learned words. A more direct way to assess and compare the contribution of PAC in relation to other factors is through conducting linear regressions, where connectivity in the learning environment (at both the phonological and semantic level), frequency and length predict the age of acquisition.

We used the frequency estimates from @braginsky2016 where unigram counts were derived based on CHILDES corpora in each language. For each word, counts included words that shared the same stem  (e.g., cats counts as cat), or words that were synonymous (e.g. father counts as daddy). For word length, we used our generated IPA transcription.

We conducted two analyses. We fit a linear regression for each language, and we fit a linear mixed-effect to all the data pooled across languages, with language as a random effect. Figure \ref{fig:regressions_img} shows the coefficient estimate for each predictor in each language, and figure \ref{fig:regressions_all_img} shows the coefficient estimates for all languages combined (all predictors were centered and scaled). The findings were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition.  Word length predicts learning in some languages such as Croatian and Norwegian, but not in others (including English). It remains, however, a significant predictor in the global model. As for the factors of interest, i.e., semantic and phonological connectivity, we also found cross-linguistic differences. The phonological connectivity contributes to learning in languages such as Croatian, English and Russian, whereas semantic connectivity contributes to learning in Turkish, Spanish and to some extent in Croatian, but interestingly not in English\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included both frequency and length as covariates. Another important difference is the number of words tested: our study uses a larger set of nouns.}. Despite this cross-linguistic variation, both phonological and semantic connectivity remain significant predictors in the global model.

@@

The present study provided a comprehensive analysis of how lexical connectivity influences the age of acquisition of nouns in toddlers. We compared two network growth scenarios and assessed the relative contribution of phonological and semantic information in 8 languages. Part of the findings largely replicate the results obtained in @hills2009, i.e., semantic networks (based on free associations) grow by preferential acquisition, not by preferential attachment. Another finding was that phonological networks also grow primarily by preferential acquisition, especially when both scenarios (PAT and PAC) were pitted against each other in the same model. These findings generalize well across languages. Moreover, both semantic and phonological connectivity in the learning environment (i.e., PAC) predict growth in a consistent way across many languages. However, when pitted against other known predictors of age of acquisition (word frequency and length), the effect of word connectivity shows a cross-linguistic variation, predicting learning in some languages, but not in others. Despite this cross-linguistic variation, both phonological and semantic connectivity contribute to the overall learning (when data is pooled across languages).

@@

```{r regressions_all_img, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, fig.cap = "\\label{fig:regressions_all_img}Estimates of predictor coefficients in the combined model pooling across languages . Ranges indicate 95\\% confidence intervals. Fade color indicates estimates of PAC predictors in a model that does not include frequency and length as covariates."}

@@

A major results of the study is that children start by learning words that have high phonological and semantic similarity with a variety of other words in the learning environment, not in the child's available lexicon.  This suggests that children are sensitive to connectivity even without having first acquired the connected words. How can children indirectly detect highly connected words, and why would such words be more readily learned? In the semantic case, free association can be predicted through the patterns of word co-occurrence [@griffiths07], meaning that highly connected words tend to be the words that co-occur with many other words in various contexts. One possibility, suggested by @hills2010, is that the referents of such words are more easily disambiguated from other potential referents. Such effect was demonstrated by cross-situational learning experiments [@smith2008].

>>>>

---
title: "Word Learning as Network Growth: A Cross-linguistic Analysis"

@@

header-includes:
   - \usepackage[subtle]{savetrees}

@@

    "Children tend to produce words earlier when they are connected to other words along both the phonological and semantic dimensions. Though this connectivity effect has been extensively documented, little is known about the underlying developmental mechanism.  One view suggests that learning is primarily driven by a network growth model where highly connected words in the child's ealy lexicon attract similar words. Another view suggests that learning is driven, not by highly connected words in the early internal lexicon, but by words that are highly connected in the external learning environment. The present study tests both scenarios systematically in both the phonological and the semantic domains, and across 8 languages. We show that external connectivity in the learning environment drives growth in both the semantic and the phonological networks, and that this pattern is consistent cross-linguistically. The findings suggest a word learning mechanism where children harness their statistical learning abilities to (indirectly) detect and learn highly connected words in the learning environment."

@@

```{r global_options, include=FALSE}
rm(list=ls())

@@

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(ggthemes)
library(xtable)
options(xtable.comment = FALSE)
library(purrr)
library(readr)
library(ggplot2)
library(langcog)
library(boot)
#library(lazyeval)
library(dplyr)
library(tidyr)
library(wordbankr)
library(directlabels)
#library(scales)
library(stringr)
library(lmtest)
library(rwebppl)
library(jsonlite)
library(nlme)
library(feather)
library(broom)
library(HDInterval)
library(BBmisc)
```

@@

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analysing dense corpora capturing language learning in context [e.g., @roy2015]. One strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn. For example, within a lexical category, words that are more frequent in child-directed speech are acquired earlier [@goodman2008]. Other factors include word length, the mean length of utterances in which the word occurs, and concreteness [see @braginsky2016].

Besides these word-level properties, the lexical structure (that is, how words relate to each other) also influences the age of acquisition of words. The lexical structure is best characterized in terms of a network where each node represents a word in the vocabulary, and each link between two nodes represents a relationship between the corresponding pair of words.  Previous studies investigated early vocabulary networks based on different word relations such as shared semantic features, target-cue relationships in free association norms, co-occurrence in child directed speech, and phonological similarity. These studies have found that children tend to produce earlier the words that have higher neighborhood density (i.e., high connectivity in the network) both at the phonological and the semantic level [@stokes2010; @storkel2009; @hills2009; @hills2010; @engelthaler2017; @stella2017; @beckage2011].

While most studies focused on the static properties of the lexical network, a few have investigated the underlying developmental process. @steyvers2005 suggested that the effect of connectivity is the consequence of how the lexical network gets constructed in the child's mind. According to this explanation, known as Preferential Attachment (PAT), highly connected words in the child's lexicon tend to "attract" more words over time, in a rich-get-richer scenario [@barabasi99]. In other words, what predicts word learning is the *internal* connectivity in the child's early lexicon. In contrast, @hills2009 suggested that what biases the learning is not the connectivity in the child's internal lexicon but, rather, *external* connectivity in the learning environment. They called this alternative explanation Preferential Acquisition (PAC).  Figure \ref{fig:growth} shows an illustration of both growth scenarios with the same simplified network. These two proposals represent two divergent ideas about the role of lexical networks in acquisition. On the PAT proposal, network structure is a causal factor in early word learning; in contrast, on the PAC approach, network structure is not internally represented and, therefore, might be an epiphenomenon of the statistics of the linguisitc input (see discussion).

```{r growth, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=2, fig.cap = "\\label{fig:growth}Filled circles (I1-I4) represent known words (internal lexicon), and empty circles (E1 and E2) represent words that have not been learned yet (external lexicon). Black lines represent links that are relevant in each growth scenario, and gray lines represent links that are irrelevant. For PAT, each candidate node is characterized with the average degree (i.e., number of links) of the existing nodes that it would attach to. Thus, according to PAT, the node E1 is more likely to enter the lexicon first. For PAC, each candidate node is characterized with its degree in the entire network. According th PAC, the node E2 is more likely to enter the lexicon first."}

@@

Studies that investigate lexical network growth have focused on semantic networks using English data [@steyvers2005; @hills2009; @hills2010].  The novelty of the current study is that it investigates whether phonological networks, like semantic networks, grow by PAC, or if they rather grow by PAT. It also provides a systematic comparison of both network growth scenarios in the phonological and the semantic domains and assesses their relative contribution to the learning process. Moreover, it tests the generality of the findings across eight languages.

<!--The paper is organized as follows. First, we describe the dataset we used, the procedure we followed to construct both semantic and phonological networks. Next, we show the results of the various analyses we conducted comparing network growth scenarios across languages. Then we test the overall contribution of the networks in the learning process relative to other known predictors of age of acquisition (frequency and length). Finally we discuss the major results in relation to known experimental facts in language development, and we speculate about the potential learning mechanism that might drive early vocabulary learning.-->

@@

plot_data_all <- ggplot(data_long, aes(x=value, y=AoA))+
  facet_grid(predictor ~ language)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=correlations, aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("AoA")

@@

plot_dist <- ggplot(data = degreeDist,  aes(x=x, y=y, col=dimension))+
  facet_grid(. ~ language)+
  geom_point(size = 0.5, alpha = 0.5) +
  scale_y_log10() + scale_x_log10() +
  theme_few()+
  theme(aspect.ratio = 0.7)+
  xlab("degree") +ylab("probability")

@@

# Networks
## Data
We used data from Wordbank [@frank2017], an open repository aggregating cross-linguistic language developmental data of the MacArthur-Bates Communicative Development Inventory (CDI), a parent report vocabulary checklist. Parent report is a reliable and valid measure of children's vocabulary that allows for the cost-effective collection of datasets large enough to test network-based models of acquisition [@fenson94].
We used the *Words and Sentence* version of the CDI which contains the productive vocabulary of toddlers (age varied between 16 to 36 months). Following previous studies [e.g., @storkel2009; @hills2009], we restricted our analysis to the category of nouns. The age of acquisition was defined by the month at which a word was produced by at least 50% of children [@goodman2008].

We obtained these nouns in eight languages: Croatian, Danish, English, Italian, Norwegian, Russian, Spanish, and Turkish. We used the subset of nouns that had entries in the Florida Association Norms (see below). Since these norms are available only in English, we used the hand-checked translation equivalents provided by Braginsky by @braginsky2016, allowing us to use the English association norms across languages. Table \ref{tab:stats} gives an overview of the data used. Translation equivalents were originally constructed for a subset of words appearing on the toddler CDI form (the subset is labeled *Words and Gestures*), and so not all words are currently available. Note, however, that all languages have at least 60% of nouns translated.

@@

tab <- xtable::xtable(statistics,
                      caption = "\\label{tab:stats}Total number of nouns produced by toddlers in the CDI (left). We included in our study the subset of these nouns that had available English translations (middle). The final set consisted of nouns that had both available translations as well entries in the Free Association Norms (right).")

@@

@hills2009, we used as an index of semantic relatedness the Florida Free Association Norms [@nelson1998]. This dataset was collected by giving adult participants a word (the cue), and asking them to write the first word that comes to mind (the target). For example, when given the word "ball", they might answer with the word "game". A pair of nodes were connected by a directed link from the cue to the target if there was a cue-target relationship between these nodes in the association norms. Each node was characterized by its "indegree", which represents the number of links for which the word was the target. To model growth from month to month, we constructed a different network at each month, made of words that have been learned at that month.

@@

We generated approximate International Phonetic Alphabet (IPA) transcriptions from the orthographic transcription, across languages, using the open source text-to-speech software **[Espeak](http://http://espeak.sourceforge.net/).** We used the Levenshtein (edit) distance as a measure of phonological relatedness between two nodes. The measure counts the minimum number of operations (insertions, deletions, substitutions) required to change one string into another.

In previous studies, two nodes were linked if they had an edit distance of 1 [@stokes2010; @storkel2009]. However, in these previous studies the network was built using an adult vocabulary. However, since the children's vocabulary contains very few word pairs with an edit distance of 1, the resulting networks were too sparse and uninformative. Thus, we increased the threshold from 1 to 2, that is, two nodes were related if their edit distance was equal to 1 or 2. Each node was characterized with its degree, i.e., the number of links it shares with other words.

@@

## Static properties of the global network

We start by analysing word connectivity in the global (static) network. We constructed this network using nouns at the oldest age for which we have CDI data (e.g., in English this corresponds to the network by 30 months). This global netowrk is the end-state towards which both PAT and PAC should converge by the last month of learning. Morover, following @hills2009, we used this end-state netowrk as a proxy for the external connectivity in the learning environment. Below we analyse properties of this global networks that are relevant to PAT and/or PAC.

###Connectivity predicts the age of acquisition

Connectivity in the global network is directly related to PAC as it represents the explicit criterion PAC uses to determine what words should be learned first (Figure \ref{fig:growth}). Therefore, a direct consequence of a PAC-like growth scenario is a correlation betweeen connectivity in the global network and the age of acquisition.\footnote{Note that this correlation is also compatible with PAT, even if this growth scenario does not rely explicity on external connectivity. Indeed, from a PAT perspective, higher connectivity in the global network is caused by earlier learning, i.e., some words end up being highly connected in the global network precisely because they happen to be acquired earlier (and therefore have a higher chance of forming more links over time).} Figure \ref{fig:data_all} shows how the age of acquisition for each word varies as a function of its connectivity/degree (or indegree for the semantic network). For ease of visual comparison, the predictor (i.e., the degree) was centered and scaled across languages. The plots show, overall, a negative correlation between the month of acquisition and the degree, indicating that nouns with higher degrees are generally learned earlier.

###Power law degree distribution?
We also analysed the global network's degree distribution. This property is particularly relevant to PAT as this growth scenario is known to generate networks with a power-law degree distribution [@barabasi99]. If the network displays this property, this fact would strongly suggests a PAT-like generative process. Conversely, if the degree distribution does not follow a power law, this fact would weaken the case for PAT. The results are shown in Figure \ref{fig:degree_distribution}. Overall, we did not find consistent evidence for a power law distribution: the extent to which the degree distribution approximates a power law varies across dimensions and languages.

For a more direct test of the developmental process, we fit explicit growth models to the data (next section).
<!--
```{r }

@@

To test the network growth scenarios, we fit different growth models to the data. We proceeded as follows. We calculated the probability that a word $w_i$, with a growth value $d_i$ would enter the lexicon at a given month, using a softmax function: 
\begin{equation}
 p(w_i)= \frac{e^{\beta d_i}}{\sum_j e^{\beta d_j} }
\end{equation}

\noindent where $\beta$ is a fitted parameter that captures the magnitude of the relationship between network parameters and growth (analogous to a regression coefficient). A positive value of $\beta$ means that words with higher growth values $d_i$ are acquired first, and a negative value means that words with lower growth values are acquired first (see Figure \ref{fig:growth} for illustrations of growth scenarios, and how a word's growth value $d_i$ is defined in each of these scenarios). The normalization includes all words that could be learned at that month.

We estimated the parameter $\beta$ using a Bayesian approach. The inference was performed using the probabilistic programming language  WebPPL [@dippl]. We defined a uniform prior over $\beta$, and at each month, we computed the likelihood function over words that could possibly enter the lexicon at that month, fit to the words that have been actually learned at that month (using formula 1). Markov chain monte carlo sampling resulted in a posterior distribution over $\beta$, which we summarized in Figure \ref{fig:pred_ind}.

Besides fitting a growth model to the data, we conducted another evaluation. This second evaluation consists in determining, in each month, the growth value distribution of all words that could possibly be learned at this month, and then computing the z-score of each learned word with respect to this distribution.  For each growth scenario, we tested if the distribution constituted by the z-scores of all learned words was different from zero, using a one-sample t-test.

The results from both evaluations were very similar and lead essentially to the same conclusions.\footnote{we do not show here the results of the second evaluation because they were redundant with the results of the first evaluation} For the semantic networks, the results replicate Hills et al.'s finding in English, which was that the semantic network grows by PAC, not by PAT. Moreover, this finding held in seven of the eight languages we examined. The PAC model also fit better than PAT for phonological networks. We note however that PAT, though weaker, fares  better for the phonological networks (where it predicts part of the growth process in some languages such as Croatian, English, Norwegian and Russian) than it does for the semantic networks (where it is rather universally unpredictive).

@@

Above we evaluated the network growth scenarios individually. As a next step, we analysed their relative contribution to the learning process. This was done through adding more fitting parameters to the model, that is, by substituting $\beta d_i$ in formula (1) with: $$\beta_{1} d_{i, 1} + \beta_{2} d_{i, 2} + \beta_{3} d_{i, 3} + \beta_{4} d_{i, 4}$$ where the indices represent the 4 networks: semPAT, semPAC, phonoPAT and PhonoPAC. Using the same fitting technique, we obtained the values shown in Figure \ref{fig:pred_ind}. PAC dominates the learning. Both phonological and semantic networks contribute to lexical growth, but the phonological network appears to be stronger and more consistent across languages. In summary, the findings show that both semantic and phonological networks contribute to the learning process, and that they both grow primarily by PAC, relying on the external connectivity in the learning environment, rather than the internal connectivity in the acquired lexicon.

@@

We saw that the way semantic and phonological information is structured in the learning environment (i.e., PAC) contribute to noun learning across languages. However, we know that other factors influence learning as well [e.g., @braginsky2016]. Next we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which correlates with phonological connectivity. <!-- [@pisoni1985].-->

Since PAT was uninformative, we dropped it from this analysis. Thus, we no longer needed to fit the growth model month-by-month as in the previous section. In fact, word utilities (i.e., growth values) in the case of PAC are fixed, they do not depend on previously learned words. A more direct way to assess and compare the contribution of PAC in relation to other word-level factors is through conducting linear regressions, where connectivity in the learning environment (at both the phonological and semantic level), frequency and length predict the age of acquisition.

We used the frequency estimates from @braginsky2016 where unigram counts were derived based on CHILDES corpora in each language (note that these frequency counts are based on transcripts from independent sets of children and represent a general estimate of environmental frequency across children). For each word, counts included words that shared the same stem  (e.g., cats counts as cat), or words that were synonymous (e.g. father counts as daddy). For word length, we used our generated IPA transcription.

We conducted two analyses. We fit a linear regression for each language, and we fit a linear mixed-effect to all the data pooled across languages, with language as a random effect. Figure \ref{fig:regressions_img} shows the coefficient estimate for each predictor in each language, and Figure \ref{fig:regressions_all_img} shows the coefficient estimates for all languages combined (all predictors were centered and scaled). The findings were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition, replicating results for nouns across a variety of analyses [@goodman2008; @braginsky2016; @roy2015]. Word length predicts learning in some languages such as Croatian and Norwegian, but not in others (including English). It remains, however, a significant predictor in the global model. As for the factors of interest, i.e., semantic and phonological connectivity, we also found cross-linguistic differences. The phonological connectivity contributes to learning in languages such as Croatian, English and Russian, whereas semantic connectivity contributes to learning in Turkish, Spanish and to some extent in Croatian, but interestingly not in English.\footnote{Semantic connectivity does not explain variance in English data beyond that explained by phonological connectivity, frequency and length. This contrasts with the original finding in Hills et al. 2009. However, in this previous study, semantic connectivity was not tested in a model that included frequency, length and phonological connectivity as covariates. Another important difference is the number of words tested: our study uses a larger set of nouns.} Despite this cross-linguistic variation, both phonological and semantic connectivity remain significant predictors in the global model.

@@

The present study provided a comprehensive analysis of how lexical connectivity influences the age of acquisition of nouns in toddlers. We compared two network growth scenarios and assessed their relative contributions of phonological vs. semantic networks across eight languages. One scenario, PAT, described a "rich get richer" network growth model in which the structure of the learner's internal network determines future growth; the other, PAC, described a model in which the external, global environmental network structure determines learners' growth patterns. Part of the findings largely replicate the results obtained in @hills2009, i.e., semantic networks (based on free associations) grow by preferential acquisition, not by preferential attachment. Another finding was that phonological networks also grow primarily by preferential acquisition, especially when both scenarios (PAT and PAC) were pitted against each other in the same model. These findings generalize well across languages. Moreover, both semantic and phonological connectivity in the learning environment (i.e., PAC) predict growth in a consistent way across many languages. However, when pitted against other known predictors of age of acquisition (word frequency and length), the effect of word connectivity shows a cross-linguistic variation, predicting learning in some languages, but not in others. Despite this cross-linguistic variation, both phonological and semantic connectivity contribute to the overall learning (when data is pooled across languages).

@@

```{r regressions_all_img, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3, fig.cap = "\\label{fig:regressions_all_img}Estimates of predictor coefficients in the combined mixed-effect model with language as a random effect. Ranges indicate 95\\% confidence intervals. Lighter points indicate estimates of PAC predictors in a model that does not include frequency and length as covariates."}

@@

One important result of the study is that children start by learning words that have high phonological and semantic similarity with a variety of other words in the learning environment, not in the child's available lexicon.  This suggests that children are sensitive to connectivity even without having first acquired the connected words. How can children indirectly detect highly connected words, and why would such words be more readily learned?
